Clustering in a hyperbolic model of complex networks.

### 2019-03-25 concentration.tex
processed most of Nikoloas comments for sections 1-3

\subsection{Joint neighborhoods and degrees in $\Gbox$}

To prove Proposition~\ref{prop:concentration_tilde_T_P_n} we need to understand the joint degree distribution in $\Gbox$. This subsequently requires us to compute the joint neighborhoods in $\Gbox$ of two points $p, p^\prime \in \Rcal$. We perform the analysis in this section. We start with a general result for near independent Poisson random variables.

@@

To make this more precise, we define, for any $0 < \varepsilon < 1$, the following set
\begin{equation}\label{eq:def_joint_degree_set_E_growing_k}
	\mathcal{E}_{\varepsilon}(k_n) = \left\{(p,p^\prime) \in \Rcal \times \Rcal
		\, : \, y,y^\prime \in \Kcal_{C}(k_n) \text{ and } |x - x^\prime| > k_n^{1 + \varepsilon} \right\}. 
\end{equation}
We will show (see Corollary~\ref{cor:expected_common_neighbors_Ecal_set}) that for all $(p,p^\prime) \in \mathcal{E}_\varepsilon(k_n)$ it holds that $\Mu{\BallPon{p} \cap \BallPon{p^\prime}} = \bigO{k_n^{1-\varepsilon}}$ and hence the joint degree distribution factorizes on this set. We will use this set later in Section~\ref{ssec:concentration_tilde_T} to prove Proposition~\ref{prop:concentration_tilde_T_P_n}. The main idea behind the above result is that if $p$ and $p^\prime$ are sufficiently separated in the $x$-direction, then the overlap of their neighborhoods $\BallPon{p} \cap \BallPon{p^\prime}$ is of smaller order than $\Mu{\BallPon{p}} + \Mu{\BallPon{p^\prime}}$. We shall therefore proceed with analyzing the joint neighborhoods in $\Gbox$.
%
%
%The following result shows that for any two points in the set
%
%The two main results, which will be the crucial technical ingredients for the proof of Proposition~\ref{prop:concentration_tilde_T_P_n}, are Lemma~\ref{lem:joint_degree_factorization} and Lemma~\ref{lem:joint_degree_distribution_shift}.

\subsubsection*{Joint neighborhoods}

Let $p, p^\prime \in \Rcal$ and denote by
%Then we denote by $\Ncal_{\text{box}}(p\Delta p^\prime)$ the number of disjoint neighbors of $p$ and $p^\prime$ in $\Gbox$, i.e. those points that belong to either $\BallPon{p}$ or $\BallPon{p^\prime}$ but not to their intersection. In addition we denote by 
$\Ncal_{\text{box}}(p,p^\prime)$ the number of joint neighbors of $p$ and $p^\prime$. We shall establish an upper bound on the expected number of joint neighbors when $p$ and $p6\prime$ are sufficiently separated. Observe that $\Exp{\Ncal_{\text{box}}(p,p^\prime)} = \Mu{\BallPon{p} \cap \BallPon{p^\prime}}$.

%as well as an asymptotic expression for the expected number points in the joint neighborhood. For this we will distinguish between the cases where the distance between the $x$-coordinates of $p$ and $p^\prime$ is small or large. Figure~\ref{fig:representation_disjoint_neighborhoods_P_n}-\ref{fig:representation_disjoint_neighborhoods_P_n_3} show the different situations that occur.

@@

%\caption{Schematic representation of the neighborhoods of $p$ and $p^\prime$ in $\Gbox$ when $|x-x^\prime| \le e^{\frac{y + y^\prime}{2}}$ used for the proof of Lemma~\ref{lem:disjoint_neighbors_P_n} and Lemma~\ref{lem:common_neighbors_Pcal_n}.}
%\label{fig:representation_disjoint_neighborhoods_P_n}

@@

%\caption{Schematic representation of the neighborhoods of $p$ and $p^\prime$ in $\Gbox$ when $e^{\frac{y + y^\prime}{2}} < |x - x^\prime| \le e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}$ used for the proof of Lemma~\ref{lem:disjoint_neighbors_P_n} and Lemma~\ref{lem:common_neighbors_Pcal_n}.}
%\label{fig:representation_disjoint_neighborhoods_P_n_2}

@@

\end{tikzpicture}
\caption{Schematic representation of the neighborhoods of $p$ and $p^\prime$ in $\Gbox$ when $|x - x^\prime| > e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}$ used for the proof of Lemma \ref{lem:disjoint_neighbors_P_n_large}. Note that although here $p^\prime \notin \BallPon{p}$, this is not true in general. This situation was merely chosen to improve readability of the figure.}
\label{fig:representation_disjoint_neighborhoods_P_n_3}
\end{figure}

We start by analyzing the shape of joint neighborhood. Due to symmetry and the fact that we have identified the left and right boundaries of the box $\Rcal$, we can, without loss of generality, assume that $p = (0,y)$ and $p^\prime = (x^\prime,y^\prime)$ with $x^\prime > 0$. To understand the computation it is helpful to have a picture. Figure~\ref{fig:representation_disjoint_neighborhoods_P_n_3} shows such an example. There are several different quantities that are important. The first are the heights where the left and right boundaries of the ball $\BallPon{p}$ go outside the box $\Rcal$. Since $x = 0$ these heights are the same and we denote this by $h(y)$. We also need to know the coordinates $\hat{y}_{\text{right}}(p,p^\prime)$ and $\hat{x}_{\text{right}}(p,p^\prime)$ of the intersection of the right boundary of the neighborhood of $p$ with the left boundary of the neighborhood of $p^\prime$ and those for the intersection of the left boundary of the neighborhood of $p$ with the right boundary of the neighborhood of $p^\prime$, which we denote by $\hat{y}_{\text{left}}(p,p^\prime)$ and $\hat{x}_{\text{left}}(p,p^\prime)$. Finally we will denote by $d(p,p^\prime)$ the distance between the lower right boundary of $\BallPon{p}$ and the lower left of $\BallPon{p^\prime}$, which is positive only when the bottom parts of both neighborhoods do not intersect, as is the case in Figure~\ref{fig:representation_disjoint_neighborhoods_P_n_3}. The condition $d(p,p^\prime) > 0$ is exactly the right notion for $p$ and $p^\prime$ being sufficiently separated.

@@

%\begin{lemma}\label{lem:disjoint_neighbors_P_n}

@@

%$\BallPon{p^\prime}$ up to the height $h^\ast := \min\{h(y), h_2(p^\prime)\}$, see Figure~\ref{fig:representation_disjoint_neighborhoods_P_n} and Figure~\ref{fig:representation_disjoint_neighborhoods_P_n_2} for reference. Note that here we do not need to consider different cases depending on whether $|x - x^\prime| \le e^{(y + y^\prime)/2}$ or $|x - x^\prime| > e^{(y + y^\prime)/2}$. 

@@

%\begin{lemma}\label{lem:disjoint_neighbors_P_n_large}

@@

%We shall compute the integral on the right. We refer to Figure~\ref{fig:representation_disjoint_neighborhoods_P_n_3} for further clarification. 

@@

%Before we proceed we show that the neighborhoods of $p$ and $p^\prime$ below $\hat{y}_{\text{right}}(p,p^\prime)$ are disjoint. This is clearly true when $\hat{y}_{\text{right}}(p,p^\prime) \le h_2(p^\prime)$ so suppose that $\hat{y}_{\text{right}}(p,p^\prime) > h_2(p^\prime)$. Then, because we identified the right and left boundaries of the box $\Rcal$ the right boundary of $\BallPon{p^\prime}$ continues from the left boundary of the box and is described by the equation

@@

%and hence the neighborhoods of $p$ and $p^\prime$ below $\hat{y}_{\text{right}}(p,p^\prime)$ are disjoint. It then follows that

@@

%\begin{figure}
%\centering
%\begin{tikzpicture}
%	
%	%The box \Rcal
%	\draw (-5,0) -- (10,0);
%	\draw (-5,4) -- (10,4);
%	
%	
%	\draw[domain=1:5,smooth,variable=\x,black] plot (\x, {2*ln(\x)});
%    \draw[domain=-1:-5,smooth,variable=\x,black] plot (\x, {2*ln(-\x)});
%    \draw[domain=6:10,smooth,variable=\x,black] plot (\x, {2*ln(\x-5)});
%    \draw[domain=0:4,smooth,variable=\x,black] plot (\x, {2*ln(5-\x)});
%    
%    
%    
%    \draw node at (-1,2.5) {$\BallPo{p} \setminus B_{\Pcal}(p^\prime)$};
%    
%    \draw node at (-0.2,3.6) {$x_1 = x^\prime - e^{\frac{y^\prime+y_1}{2}}$};
%    
%    \draw node at (2.5,3) {$\Ncal_{\Pcal}(p,p^\prime)$};
%    
%    \draw node at (5.2,3.6) {$x_2 = x+e^{\frac{y+y_2}{2}}$};
%    
%    \draw node at (6,2.5) {$B_{\Pcal}(p^\prime) \setminus \BallPo{p}$};
%    
%    \draw[dashed,thick] (2.5,1.8326) -- (2.5,0.7);
%    \draw node at (2.5,0.4) {$w(p,p^\prime)$};
%    
%    \draw[dashed,thick] (-3,1.8326) -- (8,1.8326);
%    \draw node at (-4,1.8326) {$h(p,p^\prime)$};
%    
%    \draw [decorate,decoration={brace},line width=1pt] (4,-0.1) -- (1,-0.1);
%    \draw node at (2.5,-0.45) {$d(p,p^\prime)$};
%    
%    \draw node[fill, circle, inner sep=0pt, minimum size=5pt] at (0,0) {};
%    \draw node at (0,-0.5) {$p$};
%    
%    \draw node[fill, circle, inner sep=0pt, minimum size=5pt] at (5,0) {};
%    \draw node at (5,-0.5) {$p^\prime$};
%\end{tikzpicture}
%\caption{Schematic representation of the neighborhoods of $p$ and $p^\prime$ in $\Gbox$ used for the proof of Lemma \ref{lem:common_neighbors_Pcal_n}.}
%\label{fig:representation_joint_neighborhoods}
%\end{figure}

The following result shows that if $d(p,p^\prime) > 0$, then the expected number of common neighbors is $\smallO{\Mu{\BallPon{p}} + \Mu{\BallPon{p^\prime}}}$.

\begin{lemma}\label{lem:common_neighbors_Pcal_n}
Let $p, p^\prime \in \Rcal$. Then, whenever $|x - x^\prime| > \left(e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}\right)$,
\[
	\Exp{\Ncal_{\text{box}}(p,p^\prime)} \le \left(\Mu{\BallPon{p}} + \Mu{\BallPon{p^\prime}}\right)\phi_n(p,p^\prime),
\]
where
\[
	\phi_n(p,p^\prime) = 2\left(\frac{|x - x^\prime|}{e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}}\right)^{-(2\alpha - 1)} 
	+ \frac{3 \nu^{2\alpha + 1}e^{-(\alpha - \frac{1}{2})R} e^{\alpha y}}{2 \pi^{2\alpha}\left(
	e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}\right)}
	+ \frac{\nu e^{-(\alpha - \frac{1}{2})R}}{e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}}. 
\]
\end{lemma}

@@

We now return to the joint degree distribution of nodes in $\Gbox$. Recall the definition of $\mathcal{E}_\varepsilon(k_n)$
\[
	\mathcal{E}_{\varepsilon}(k_n) = \left\{(p,p^\prime) \in \Rcal \times \Rcal
			\, : \, y,y^\prime \in \Kcal_{C}(k_n) \text{ and } |x - x^\prime| > k_n^{1 + \varepsilon} \right\}. 
\]
The following result, which follows from Lemma~\ref{lem:common_neighbors_Pcal_n}, shows that on this set, the expected number of common neighbors is $\smallO{k_n}$.

\begin{lemma}\label{cor:expected_common_neighbors_Ecal_set}
Fix $0 < \varepsilon < 1$ and let $\varepsilon^\prime = \min\{\varepsilon(2\alpha - 1),\varepsilon\}$. Then for all $(p,p^\prime) \in \mathcal{E}_\varepsilon(k_n)$, as $n \to \infty$,
\[
	\Mu{\BallPon{p} \cup \BallPon{p^\prime}} = \bigO{k_n^{1-\varepsilon^\prime}}.
\] 
\end{lemma}

\begin{proof}
Since for all $(p,p^\prime) \in \mathcal{E}_\varepsilon(k_n)$ we have $\Mu{\BallPon{p}}, \Mu{\BallPon{p^\prime}} = \bigT{k_n}$, Lemma~\ref{lem:common_neighbors_Pcal_n} implies that
\[
	\Mu{\BallPon{p} \cup \BallPon{p^\prime}} \le \bigO{k_n} \phi_n(p,p^\prime),
\]
where
\[
	\phi_n(p,p^\prime) = 2\left(\frac{|x - x^\prime|}{e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}}\right)^{-(2\alpha - 1)} 
		+ \frac{3 \nu^{2\alpha + 1}e^{-(\alpha - \frac{1}{2})R} e^{\alpha y}}{2 \pi^{2\alpha}\left(
		e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}\right)}
		+ \frac{\nu e^{-(\alpha - \frac{1}{2})R}}{e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}}. 
\]
We thus need to show that $\phi_n(p,p^\prime) = \bigO{k_n^{-\varepsilon}}$. For $(p,p^\prime) \in \mathcal{E}_\varepsilon(k_n)$, it holds that $e^{y/2}, e^{y^\prime/2} = \bigT{k_n}$ and $|x - x^\prime| > k_n^{1+\varepsilon}$ and hence
\begin{align*}
	 2\left(\frac{|x - x^\prime|}{e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}}\right)^{-(2\alpha - 1)}
	 =	\bigO{k_n^{-\varepsilon(2\alpha - 1)}}
\end{align*}
For the second term in $\phi_n(p,p^\prime)$ we use that $e^{\alpha y^\ast} = \bigT{k_n^{2\alpha}}$ and $e^{R} = \bigT{n^2}$ to obtain
\[
	\frac{3 \nu^{2\alpha + 1}e^{-(\alpha - \frac{1}{2})R} e^{\alpha y}}{2 \pi^{2\alpha}\left(
			e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}\right)}
	= \bigO{1} n^{-(2\alpha - 1)} k_n^{2\alpha - 1} = \bigO{n^{-(\alpha - \frac{1}{2})}}.
\]
Finally, the third term satisfies $\bigO{n^{-(2\alpha - 1)}k_n^{-1}}$, and we conclude that
\[
	\phi_n(p,p^\prime) = \bigO{k_n^{-\varepsilon(2\alpha - 1)} + n^{-(\alpha - \frac{1}{2})}
	+ n^{-(2\alpha - 1)}k_n^{-1}} = \bigO{k_n^{-\varepsilon^\prime}},
\]
where we used that $\varepsilon^\prime = \min\{\varepsilon(2\alpha - 1),\varepsilon\}$. 
\end{proof}

@@

\begin{proof}
Define the random variables
\begin{align*}
	X_1(p,p^\prime) &:= \Po\left(\Mu{\BallPon{p}\setminus \BallPon{p^\prime}}\right),\\
	X_2(p,p^\prime) &:= \Po\left(\Mu{\BallPon{p^\prime}\setminus \BallPon{p}}\right),\\
	Y(p,p^\prime) &:= \Po\left(\Mu{\BallPon{p} \cup \BallPon{p^\prime}}\right),
\end{align*}
so that
\begin{align*}
	\rho_{\text{box}}(p,p^\prime,k_n-t_n,k_n-t_n^\prime) &= \Prob{X_1(p,p^\prime) + Y(p,p^\prime) = k_n-t_n, X_2(p,p^\prime) + Y(p,p^\prime) = k_n - t_n^\prime}.
\end{align*}
Since by Lemma~\ref{cor:expected_common_neighbors_Ecal_set} $\Mu{\BallPon{p} \cup \BallPon{p^\prime}} = \bigO{k_n^{1-\varepsilon^\prime}}$, it follows from Lemma~\ref{lem:near_independence_poisson} that
\[
	\rho_{\text{box}}(p,p^\prime,k_n-t_n,k_n-t_n^\prime) = (1+\smallO{1})\rho_{\text{box}}(p,k_n-t_n)\rho_{\text{box}}(p^\prime,k_n-t_n^\prime).
\]
The result then follows by applying Lemma~\ref{lem:finite_shifts_poisson} twice.
\end{proof}

>>>>

\subsection{Joint neighbourhoods and degrees in $\Gbox$}

To prove Proposition~\ref{prop:concentration_tilde_T_P_n} we need to understand the joint degree distribution in $\Gbox$. This subsequently requires us to compute the joint neighbourhoods in $\Gbox$ of two points $p, p^\prime \in \Rcal$. We perform the analysis in this section. We start with a general result for near independent Poisson random variables.

@@

To make this more precise, we define, for any $0 < \varepsilon < 1$, the following set
\begin{equation}\label{eq:def_joint_degree_set_E_growing_k}
	\mathcal{E}_{\varepsilon}(k_n) = \left\{(p,p^\prime) \in \Rcal \times \Rcal
		\, : \, y,y^\prime \in \Kcal_{C}(k_n) \text{ and } |x - x^\prime| > k_n^{1 + \varepsilon} \right\}. 
\end{equation}
We will show (see Corollary~\ref{cor:expected_common_neighbours_Ecal_set}) that for all $(p,p^\prime) \in \mathcal{E}_\varepsilon(k_n)$ it holds that $\Mu{\BallPon{p} \cap \BallPon{p^\prime}} = \bigO{k_n^{1-\varepsilon}}$ and hence the joint degree distribution factorizes on this set. We will use this set later in Section~\ref{ssec:concentration_tilde_T} to prove Proposition~\ref{prop:concentration_tilde_T_P_n}. The main idea behind the above result is that if $p$ and $p^\prime$ are sufficiently separated in the $x$-direction, then the overlap of their neighbourhoods $\BallPon{p} \cap \BallPon{p^\prime}$ is of smaller order than $\Mu{\BallPon{p}} + \Mu{\BallPon{p^\prime}}$. We shall therefore proceed with analyzing the joint neighbourhoods in $\Gbox$.
%
%
%The following result shows that for any two points in the set
%
%The two main results, which will be the crucial technical ingredients for the proof of Proposition~\ref{prop:concentration_tilde_T_P_n}, are Lemma~\ref{lem:joint_degree_factorization} and Lemma~\ref{lem:joint_degree_distribution_shift}.

\subsubsection*{Joint neighbourhoods}

Let $p, p^\prime \in \Rcal$ and denote by
%Then we denote by $\Ncal_{\text{box}}(p\Delta p^\prime)$ the number of disjoint neighbours of $p$ and $p^\prime$ in $\Gbox$, i.e. those points that belong to either $\BallPon{p}$ or $\BallPon{p^\prime}$ but not to their intersection. In addition we denote by 
$\Ncal_{\text{box}}(p,p^\prime)$ the number of joint neighbours of $p$ and $p^\prime$. We shall establish an upper bound on the expected number of joint neighbours when $p$ and $p6\prime$ are sufficiently separated. Observe that $\Exp{\Ncal_{\text{box}}(p,p^\prime)} = \Mu{\BallPon{p} \cap \BallPon{p^\prime}}$.

%as well as an asymptotic expression for the expected number points in the joint neighbourhood. For this we will distinguish between the cases where the distance between the $x$-coordinates of $p$ and $p^\prime$ is small or large. Figure~\ref{fig:representation_disjoint_neighbourhoods_P_n}-\ref{fig:representation_disjoint_neighbourhoods_P_n_3} show the different situations that occur.

@@

%\caption{Schematic representation of the neighbourhoods of $p$ and $p^\prime$ in $\Gbox$ when $|x-x^\prime| \le e^{\frac{y + y^\prime}{2}}$ used for the proof of Lemma~\ref{lem:disjoint_neighbours_P_n} and Lemma~\ref{lem:common_neighbours_Pcal_n}.}
%\label{fig:representation_disjoint_neighbourhoods_P_n}

@@

%\caption{Schematic representation of the neighbourhoods of $p$ and $p^\prime$ in $\Gbox$ when $e^{\frac{y + y^\prime}{2}} < |x - x^\prime| \le e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}$ used for the proof of Lemma~\ref{lem:disjoint_neighbours_P_n} and Lemma~\ref{lem:common_neighbours_Pcal_n}.}
%\label{fig:representation_disjoint_neighbourhoods_P_n_2}

@@

\end{tikzpicture}
\caption{Schematic representation of the neighbourhoods of $p$ and $p^\prime$ in $\Gbox$ when $|x - x^\prime| > e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}$ used for the proof of Lemma \ref{lem:disjoint_neighbours_P_n_large}. Note that although here $p^\prime \notin \BallPon{p}$, this is not true in general. This situation was merely chosen to improve readability of the figure.}
\label{fig:representation_disjoint_neighbourhoods_P_n_3}
\end{figure}

We start by analyzing the shape of joint neighbourhood. Due to symmetry and the fact that we have identified the left and right boundaries of the box $\Rcal$, we can, without loss of generality, assume that $p = (0,y)$ and $p^\prime = (x^\prime,y^\prime)$ with $x^\prime > 0$. To understand the computation it is helpful to have a picture. Figure~\ref{fig:representation_disjoint_neighbourhoods_P_n_3} shows such an example. There are several different quantities that are important. The first are the heights where the left and right boundaries of the ball $\BallPon{p}$ go outside the box $\Rcal$. Since $x = 0$ these heights are the same and we denote this by $h(y)$. We also need to know the coordinates $\hat{y}_{\text{right}}(p,p^\prime)$ and $\hat{x}_{\text{right}}(p,p^\prime)$ of the intersection of the right boundary of the neighbourhood of $p$ with the left boundary of the neighbourhood of $p^\prime$ and those for the intersection of the left boundary of the neighbourhood of $p$ with the right boundary of the neighbourhood of $p^\prime$, which we denote by $\hat{y}_{\text{left}}(p,p^\prime)$ and $\hat{x}_{\text{left}}(p,p^\prime)$. Finally we will denote by $d(p,p^\prime)$ the distance between the lower right boundary of $\BallPon{p}$ and the lower left of $\BallPon{p^\prime}$, which is positive only when the bottom parts of both neighbourhoods do not intersect, as is the case in Figure~\ref{fig:representation_disjoint_neighbourhoods_P_n_3}. The condition $d(p,p^\prime) > 0$ is exactly the right notion for $p$ and $p^\prime$ being sufficiently separated.

@@

%\begin{lemma}\label{lem:disjoint_neighbours_P_n}

@@

%$\BallPon{p^\prime}$ up to the height $h^\ast := \min\{h(y), h_2(p^\prime)\}$, see Figure~\ref{fig:representation_disjoint_neighbourhoods_P_n} and Figure~\ref{fig:representation_disjoint_neighbourhoods_P_n_2} for reference. Note that here we do not need to consider different cases depending on whether $|x - x^\prime| \le e^{(y + y^\prime)/2}$ or $|x - x^\prime| > e^{(y + y^\prime)/2}$. 

@@

%\begin{lemma}\label{lem:disjoint_neighbours_P_n_large}

@@

%We shall compute the integral on the right. We refer to Figure~\ref{fig:representation_disjoint_neighbourhoods_P_n_3} for further clarification. 

@@

%Before we proceed we show that the neighbourhoods of $p$ and $p^\prime$ below $\hat{y}_{\text{right}}(p,p^\prime)$ are disjoint. This is clearly true when $\hat{y}_{\text{right}}(p,p^\prime) \le h_2(p^\prime)$ so suppose that $\hat{y}_{\text{right}}(p,p^\prime) > h_2(p^\prime)$. Then, because we identified the right and left boundaries of the box $\Rcal$ the right boundary of $\BallPon{p^\prime}$ continues from the left boundary of the box and is described by the equation

@@

%and hence the neighbourhoods of $p$ and $p^\prime$ below $\hat{y}_{\text{right}}(p,p^\prime)$ are disjoint. It then follows that

@@

%\begin{figure}
%\centering
%\begin{tikzpicture}
%	
%	%The box \Rcal
%	\draw (-5,0) -- (10,0);
%	\draw (-5,4) -- (10,4);
%	
%	
%	\draw[domain=1:5,smooth,variable=\x,black] plot (\x, {2*ln(\x)});
%    \draw[domain=-1:-5,smooth,variable=\x,black] plot (\x, {2*ln(-\x)});
%    \draw[domain=6:10,smooth,variable=\x,black] plot (\x, {2*ln(\x-5)});
%    \draw[domain=0:4,smooth,variable=\x,black] plot (\x, {2*ln(5-\x)});
%    
%    
%    
%    \draw node at (-1,2.5) {$\BallPo{p} \setminus B_{\Pcal}(p^\prime)$};
%    
%    \draw node at (-0.2,3.6) {$x_1 = x^\prime - e^{\frac{y^\prime+y_1}{2}}$};
%    
%    \draw node at (2.5,3) {$\Ncal_{\Pcal}(p,p^\prime)$};
%    
%    \draw node at (5.2,3.6) {$x_2 = x+e^{\frac{y+y_2}{2}}$};
%    
%    \draw node at (6,2.5) {$B_{\Pcal}(p^\prime) \setminus \BallPo{p}$};
%    
%    \draw[dashed,thick] (2.5,1.8326) -- (2.5,0.7);
%    \draw node at (2.5,0.4) {$w(p,p^\prime)$};
%    
%    \draw[dashed,thick] (-3,1.8326) -- (8,1.8326);
%    \draw node at (-4,1.8326) {$h(p,p^\prime)$};
%    
%    \draw [decorate,decoration={brace},line width=1pt] (4,-0.1) -- (1,-0.1);
%    \draw node at (2.5,-0.45) {$d(p,p^\prime)$};
%    
%    \draw node[fill, circle, inner sep=0pt, minimum size=5pt] at (0,0) {};
%    \draw node at (0,-0.5) {$p$};
%    
%    \draw node[fill, circle, inner sep=0pt, minimum size=5pt] at (5,0) {};
%    \draw node at (5,-0.5) {$p^\prime$};
%\end{tikzpicture}
%\caption{Schematic representation of the neighbourhoods of $p$ and $p^\prime$ in $\Gbox$ used for the proof of Lemma \ref{lem:common_neighbours_Pcal_n}.}
%\label{fig:representation_joint_neighbourhoods}
%\end{figure}

The following result shows that if $d(p,p^\prime) > 0$, then the expected number of common neighbours is $\smallO{\Mu{\BallPon{p}} + \Mu{\BallPon{p^\prime}}}$.

\begin{lemma}\label{lem:common_neighbours_Pcal_n}
Let $p, p^\prime \in \Rcal$. Then, whenever $|x - x^\prime| > \left(e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}\right)$,
\[
	\Exp{\Ncal_{\text{box}}(p,p^\prime)} \le \left(\Mu{\BallPon{p}} + \Mu{\BallPon{p^\prime}}\right)\phi_n(p,p^\prime),
\]
where
\[
	\phi_n(p,p^\prime) = 2\left(\frac{|x - x^\prime|}{e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}}\right)^{-(2\alpha - 1)} 
	+ \frac{3 \nu^{2\alpha + 1}e^{-(\alpha - \frac{1}{2})R} e^{\alpha y}}{2 \pi^{2\alpha}\left(
	e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}\right)}
	+ \frac{\nu e^{-(\alpha - \frac{1}{2})R}}{e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}}. 
\]
\end{lemma}

@@

We now return to the joint degree distribution of nodes in $\Gbox$. Recall the definition of $\mathcal{E}_\varepsilon(k_n)$
\[
	\mathcal{E}_{\varepsilon}(k_n) = \left\{(p,p^\prime) \in \Rcal \times \Rcal
			\, : \, y,y^\prime \in \Kcal_{C}(k_n) \text{ and } |x - x^\prime| > k_n^{1 + \varepsilon} \right\}. 
\]
The following result, which follows from Lemma~\ref{lem:common_neighbours_Pcal_n}, shows that on this set, the expected number of common neighbours is $\smallO{k_n}$.

\begin{lemma}\label{cor:expected_common_neighbours_Ecal_set}
Fix $0 < \varepsilon < 1$ and let $\varepsilon^\prime = \min\{\varepsilon(2\alpha - 1),\varepsilon\}$. Then for all $(p,p^\prime) \in \mathcal{E}_\varepsilon(k_n)$, as $n \to \infty$,
\[
	\Mu{\BallPon{p} \cup \BallPon{p^\prime}} = \bigO{k_n^{1-\varepsilon^\prime}}.
\] 
\end{lemma}

\begin{proof}
Since for all $(p,p^\prime) \in \mathcal{E}_\varepsilon(k_n)$ we have $\Mu{\BallPon{p}}, \Mu{\BallPon{p^\prime}} = \bigT{k_n}$, Lemma~\ref{lem:common_neighbours_Pcal_n} implies that
\[
	\Mu{\BallPon{p} \cup \BallPon{p^\prime}} \le \bigO{k_n} \phi_n(p,p^\prime),
\]
where
\[
	\phi_n(p,p^\prime) = 2\left(\frac{|x - x^\prime|}{e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}}\right)^{-(2\alpha - 1)} 
		+ \frac{3 \nu^{2\alpha + 1}e^{-(\alpha - \frac{1}{2})R} e^{\alpha y}}{2 \pi^{2\alpha}\left(
		e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}\right)}
		+ \frac{\nu e^{-(\alpha - \frac{1}{2})R}}{e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}}. 
\]
We thus need to show that $\phi_n(p,p^\prime) = \bigO{k_n^{-\varepsilon}}$. For $(p,p^\prime) \in \mathcal{E}_\varepsilon(k_n)$, it holds that $e^{y/2}, e^{y^\prime/2} = \bigT{k_n}$ and $|x - x^\prime| > k_n^{1+\varepsilon}$ and hence
\begin{align*}
	 2\left(\frac{|x - x^\prime|}{e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}}\right)^{-(2\alpha - 1)}
	 =	\bigO{k_n^{-\varepsilon(2\alpha - 1)}}
\end{align*}
For the second term in $\phi_n(p,p^\prime)$ we use that $e^{\alpha y^\ast} = \bigT{k_n^{2\alpha}}$ and $e^{R} = \bigT{n^2}$ to obtain
\[
	\frac{3 \nu^{2\alpha + 1}e^{-(\alpha - \frac{1}{2})R} e^{\alpha y}}{2 \pi^{2\alpha}\left(
			e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}\right)}
	= \bigO{1} n^{-(2\alpha - 1)} k_n^{2\alpha - 1} = \bigO{n^{-(\alpha - \frac{1}{2})}}.
\]
Finally, the third term satisfies $\bigO{n^{-(2\alpha - 1)}k_n^{-1}}$, and we conclude that
\[
	\phi_n(p,p^\prime) = \bigO{k_n^{-\varepsilon(2\alpha - 1)} + n^{-(\alpha - \frac{1}{2})}
	+ n^{-(2\alpha - 1)}k_n^{-1}} = \bigO{k_n^{-\varepsilon^\prime}},
\]
where we used that $\varepsilon^\prime = \min\{\varepsilon(2\alpha - 1),\varepsilon\}$. 
\end{proof}

@@

\begin{proof}
Define the random variables
\begin{align*}
	X_1(p,p^\prime) &:= \Po\left(\Mu{\BallPon{p}\setminus \BallPon{p^\prime}}\right),\\
	X_2(p,p^\prime) &:= \Po\left(\Mu{\BallPon{p^\prime}\setminus \BallPon{p}}\right),\\
	Y(p,p^\prime) &:= \Po\left(\Mu{\BallPon{p} \cup \BallPon{p^\prime}}\right),
\end{align*}
so that
\begin{align*}
	\rho_{\text{box}}(p,p^\prime,k_n-t_n,k_n-t_n^\prime) &= \Prob{X_1(p,p^\prime) + Y(p,p^\prime) = k_n-t_n, X_2(p,p^\prime) + Y(p,p^\prime) = k_n - t_n^\prime}.
\end{align*}
Since by Lemma~\ref{cor:expected_common_neighbours_Ecal_set} $\Mu{\BallPon{p} \cup \BallPon{p^\prime}} = \bigO{k_n^{1-\varepsilon^\prime}}$, it follows from Lemma~\ref{lem:near_independence_poisson} that
\[
	\rho_{\text{box}}(p,p^\prime,k_n-t_n,k_n-t_n^\prime) = (1+\smallO{1})\rho_{\text{box}}(p,k_n-t_n)\rho_{\text{box}}(p^\prime,k_n-t_n^\prime).
\]
The result then follows by applying Lemma~\ref{lem:finite_shifts_poisson} twice.
\end{proof}

### 2019-03-25 concentration_argument.tex
processed most of Nikoloas comments for sections 1-3

Although powerful, the current versions of the concentration of heights arguments are only valid for the function $\rho(y,k_n) := \Prob{\Po\left(\Mu{\BallPo{y}}\right) = k_n}$, which uses the neighborhoods in the infinite model $\Ginf$. Since we will also be working in the Poissonized KPKVB model $\GPo$ and the finite box model $\Gbox$, we would like to use concentration of heights arguments when the integrand contains either the function $\rho_{\Po}(y,k_n)$ or $\rho_{\text{box}}(y,k_n)$. To establish these result we first prove a more general version of Lemma~\ref{lem:concentration_argument}.

>>>>

Although powerful, the current versions of the concentration of heights arguments are only valid for the function $\rho(y,k_n) := \Prob{\Po\left(\Mu{\BallPo{y}}\right) = k_n}$, which uses the neighbourhoods in the infinite model $\Ginf$. Since we will also be working in the Poissonized KPKVB model $\GPo$ and the finite box model $\Gbox$, we would like to use concentration of heights arguments when the integrand contains either the function $\rho_{\Po}(y,k_n)$ or $\rho_{\text{box}}(y,k_n)$. To establish these result we first prove a more general version of Lemma~\ref{lem:concentration_argument}.
